Local PCA Shows How the Effect of Population Structure Differs Along the Genome

### 2018-10-30 writeup/review-responses.tex

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We are writing to submit a revised version of our manuscript, 
``Local PCA shows how the effect of population structure differs along the genome''.
We appreciate the thoughtful comments by the reviewers and the editors,
and have made changes to address all of the points raised.
The changes are mostly to what is discussed;
we sympathize with the additional testing suggested by Reviewer 1,
but think that it falls outside the scope of this paper,
and would be appropriate for follow-up work that builds on our descriptive method
to create more specific, hypothesis-driven tools.

We have also included a color-coded manuscript highlighting differences 
between this version and the last one (but the line numbers referred to in the 
``response to reviewers'' refer to the standard manuscript).
The file is named ``\url{local_pca_paper-diff164c6.pdf}'',
and is labeled another ``response to reviewers'' for lack of a better category in the submission system.

\vspace{5em}

@@

\begin{quote}
    I have not previously reviewed the manuscript and but I think that most of the comments of previous reviewers were well reacted upon. However, the resulting manuscript is to a new reader a bit unclear in terms of the application and usability of the method and ability to distinguish between different scenarios leading to observed patterns and I think further simulations of mechanisms of interest should be performed with the aim to either demonstrate what the limitations of the method are for potential users or to further explore biological processes hinted at.
\end{quote}

A similar point was raised by another reviewer in the previous round.
It might be reasonable, as \citet{ma2012investigation} did,
to use these results to develop an inference method for a particular signal.
We have rather chosen to follow the lead of genomic PCA,
which, despite having connections to demographic quantities,
is most widely-used as a purely descriptive tool,
for communicating population genetics results.
We've tried to make this a bit more clear in the introduction. \llname{ll:descriptive}

\begin{point}{}
    Is the method meant to facilitate discovery of new inversions? If so, what number of samples are necessary? Of what coverage (what level of missingness is allowed)? What size of inversions is possible to detect? What are ranges of the distance values between windows that could point to inversions (any significance test would be recommended)? How does the method compare to other methods that detect inversions (esp. Ma \& Amos 2012)? If it is not meant to discover new inversions and it should only confirm known existing inversions, what is the success rate (with varying sample size, missingness)?
\end{point}

\reply{
    These are excellent questions for a method designed to detect inversions.
    However, we do not set out to make such a method,
    but rather a visualization method in which, empirically, inversions often leave a large signal.
    We are not sure if it would be productive to turn this into a quantitative method, either:
    as we have further clarified in the paper \llname{ll:inversion_caveat},
    distinguishing inversions from regions of low recombination
    likely requires other sources of information.
}

@@

\begin{point}{}
    The authors argue that linked selection is the likely cause for the patterns they observed in Drosophila dataset since the recombination rate correlates with the position of the first MDS coordinate (from line number 321). At the same time, the authors mention that the ability to identify the windows could be influenced by recombination rate in the case of linked selection and also in the presence of recombination hotspots (section 3.1 Validation). As such, I do not agree with the authors, that it is enough to demonstrate that linked selection could cause these patterns but to use this in interpretation, they should be able to distinguish this explanation from the other possible causes in simulations first and then on real data (e.g., using recombination maps for Drosophila to simulate exactly the same regions could be of advantage).
\end{point}

\reply{
    We don't share the reviewer's concern about variation in recombination rate:
    although a region of low recombination could be dominated by a single tree,
    and therefore provide a strongly divergent PCA signal,
    there is no reason this would provide a \emph{consistent} signal
    across a large fraction of a chromosome,
    as we see here.
    We've tried to make this argument more clear. {\llname{ll:not_neutral}}
    More generally, we heartily agree that we would love to have available a test
    that could prove to what degree these (or any) varying patterns of relatedness along genomes
    are due to linked selection, demographic noise, or some other force.
    However, this is one of the major unsolved question in the field,
    and is subject to substantial debate.
    % (see for instance \citet{jensen2014unfounded,harris2018unfounded} and subsequent rejoinders).
    We've added more to the discussion on this. \revref
}

@@

    I would like to see evidence that the windows that are found by this method are not associated with ``which'' PCs happen to be included in the top K. To do that, you could for example calculate the top $M>K$ pcs, check the correlation with the overall PCs, and then plot ``presence'' of a PC (measured in correlation in the top K) against the MDS values. It would alternatively be possible to make an argument that it will always be fine if you use a large enough number of PCs, but this is argued to have severe computational problems (an argument I don't know that I buy). It would be enough to do this for retaining *one* PC in the simulation that has 2 \emph{equally strong} PCs, but I have a suspicion that this case goes quite wrong and is anyway unrealistic. Better to keep more, e.g. 5 when there is non-trivial structure in the truth, and see that the 4th or 5th swapping with the 6th or 7th doesn't affect much. To be clear: the authors shouldn't necessarily follow this route, but they should address the underlying concern.
\end{point}

\reply{
    Apologies for not realizing the precise nature of the original concern.
    Yes, we agree that if strongly different PCs switched order 
    right at the boundary of which PCs were included,
    then that could easily show up as (spurious) structure in our MDS plots.
    However, as the reviewer says,
    spurious structure created by this problem would dissappear when using a larger number of PCs
    -- this does not happen with our empirical results.
    We have added discussion of these points. {\revref}
    Furthermore, and speculatively,
    our method of choosing a window size looks for a big enough window that the PCs within a window 
    are relatively stable,
    and so might help prevent sensitivity to PC switching.
}

### 2018-10-30 writeup/local_pca_paper.tex
final tweaks

Patterns in genome-wide relatedness are often summarized
by applying principal components analysis (PCA, \citet{patterson2006population}) 
to the genetic data matrix,
as inspired by the pioneering work of \citet{menozzi1978synthetic}. \revpoint{AE}{3}
% About 37 years ago, \citet{menozzi1978synthetic} first applied principal component analysis (PCA) in population genetics to construct maps summarizing genetic variation \citep{menozzi1978synthetic}. 
% Nowadays, PCA is a widely used powerful non-parametric method to extract information from genetic data.
The results of PCA can be related to the genealogical history of the samples, 
such as time to most recent common ancestor and migration rate between populations \citep{novembre2008interpreting,mcvean2009genealogical}, 
and sometimes produce ``maps'' of population structure
that reflect the samples' geographic origin distorted by rates of gene flow
\citep{novembre2008genes}.
% although other patterns can be caused by recent migration or nongeographic kinship patterns.  % \citep{astle2009population}.
% Through dimension-reduction, PCA can identify key components of population structure, which describes how different samples are related, and are often closely related to geography. 
% Plots of the first two principal components (PCs) can mimic the samples' geographic origin to some extent. 
% Since population structure describes how different samples are related, samples living closer tend to be more genetically similar and thus tend to be clustered in PC plots \citep{novembre2008genes,patterson2006population}. 
% However this relatedness is limited while there's recent migration or for group with nongeographic kinship patterns,for example, social or religious groups. \citep{astle2009population}

@@

Simulations with linked selection produced mixed results
(\SFigure \ref{sfig:slim_sims2}).
The method strongly identified the regions under spatially varying linked selection
(also our simulation with strongest positive selection). \revpoint{2}{2}
It also identified the regions (although less unambiguously)
with constant selection and stepwise varying recombination rate,
but did not clearly identify them with constant recombination rate.
This difference is likely because recombination rates are overall lower in the first case,
leading to a stronger effect of linked selection.
These tests are not meant to be comprehensive survey of linked selection,
but only to demonstrate that linked selection can produce signals similar to what we see in real data.

>>>>

Patterns in genome-wide relatedness are often summarized
by applying principal components analysis (PCA, \citet{patterson2006population}) 
to the genotype matrix,
as inspired by the pioneering work of \citet{menozzi1978synthetic}. \revpoint{AE}{3}
% About 37 years ago, \citet{menozzi1978synthetic} first applied principal component analysis (PCA) in population genetics to construct maps summarizing genetic variation \citep{menozzi1978synthetic}. 
% Nowadays, PCA is a widely used powerful non-parametric method to extract information from genetic data.
The results of PCA can be related to the genealogical history of the samples, 
such as time to most recent common ancestor and migration rate between populations \citep{novembre2008interpreting,mcvean2009genealogical}, 
and sometimes produce ``maps'' of population structure
that reflect the samples' geographic origin distorted by rates of gene flow
\citep{novembre2008genes}.
% although other patterns can be caused by recent migration or nongeographic kinship patterns.  % \citep{astle2009population}.
% Through dimension-reduction, PCA can identify key components of population structure, which describes how different samples are related, and are often closely related to geography. 
% Plots of the first two principal components (PCs) can mimic the samples' geographic origin to some extent. 
% Since population structure describes how different samples are related, samples living closer tend to be more genetically similar and thus tend to be clustered in PC plots \citep{novembre2008genes,patterson2006population}. 
% However this relatedness is limited while there's recent migration or for group with nongeographic kinship patterns,for example, social or religious groups. \citep{astle2009population}

@@

Simulations with linked selection produced mixed results
(\SFigure \ref{sfig:slim_sims2}).
The method strongly identified the distinct regions 
in the simulation with spatially varying linked selection.
It also identified the regions (although less unambiguously)
with constant selection and stepwise varying recombination rate,
but did not clearly identify them with constant recombination rate.
The differences in power between these three scenarios are likely
explained by varying strength of linked selection -- 
the simulation with spatially varying selection was also the case with strongest positive selection,
and recombination rates were overall lower in the simulation with stepwise varying recombination rate
than in the simulation with constant rate. \revpoint{2}{2}
These tests are not meant to be comprehensive survey of linked selection,
but only to demonstrate that linked selection can produce signals similar to what we see in real data.
